A Simple Type Theory for Category Theory

### 2022-02-13 paper/lipics-v2021-sample-article.tex
some semantics

In this section we establish our desired semantic soundness theorem,
that all constructions in the theory can be interpreted in any virtual
equipment with the corresponding universal properties.
%
This proof is straightforward, since our constructions are explicitly
modeled on the universal properties, with the main difference being
that our notion of restriction is strict.

\begin{definition}[Psuedo Double Category]

@@

\begin{definition}[Restriction]

\begin{definition}[Unit]

>>>>

In this section we establish our desired semantic soundness and
completeness theorems for models in virtual equipments with the
corresponding universal properties.
%
The first component of this is \emph{soundness}: every category,
object, set and transformation denote an object, vertical arrow,
horizontal arrow and 2-cell, respectively in any representable virtual
equipment, such that all equations in the theory are satisfied.
%
We combine this with a set theoretic semantics for types $A$ and
contexts $\Gamma$ to get a simple semantics of our logic over a
virtual equipment.
%
The second component of this is the \emph{completeness} theorem:
informally, the theory contains every category, object, set and
transformation that \emph{must} exist in any virtual equipment, and
further equations between syntactic forms hold if and only if they
hold in every model.
%
While sounding more ambitious, the completeness theorem is
typically simpler as it just amounts to showing that the syntax itself
forms a virtual equipment.

\subsection{Notion of Model}

\begin{definition}[Virtual Double Category]
  A virtual double category $V$ consists of
  \begin{enumerate}
    \item A set $V_0$ of objects.
    \item For each $\cat C, \cat D \in V_0$, a set $V_v(\cat C, \cat
      D)$ of vertical arrows, with composition and unit operations
      satisfying the category axioms.
    \item A set $V_h$ of horizontal arrows, with $d^+,d^- : V_h \to
      V_0$.
    \item For each composable string $\Phi$ of horizontal arrows
      starting at $\cat C$ and ending at $\cat D$, and horizontal
      arrow $R$, a set $V_2(\Phi; R)$ of 2-cells, with projection
      functions $d^{\pm} : V_2(\Phi;R) \to V_v(d^\pm \Phi, d^\pm R)$.

    \item For each $R \in V_h$, an identity arrow $\id_R \in
      V_2(d^-R,R,d^+ R; R)$ with $d^\pm \id_R = \id_{d^\pm R}$.

    \item For any $2$-cell $\alpha \in V_2(\Phi;R)$ and composable
      string $\phi$ of 2-cells into $\Phi$, a composite $\alpha \circ
      \phi$.

    \item Composition should be unital in that:
      \[ \id_R \circ [d^-\alpha, \alpha, d^+\alpha] = \alpha  \]
      and
      \[ \alpha \circ [\vec \id] = \alpha\]
    \item And associative in that
      \[ (\alpha \circ [\beta_1,\ldots]) \circ [\gamma_{11},\ldots,\gamma_{21},\ldots,] = \alpha \circ [(\beta_1 \circ \gamma_{11},\ldots), (\beta_2 \circ \gamma_{21},\ldots)]\]
  \end{enumerate}
\end{definition}

\begin{definition}[Restriction]
  Let $R \in V_h$ and $f \in V_v(\cat C, d^-R)$ and $g \in V_v(\cat D,
  d^+R)$, then a restriction of $R$ along $f,g$ is a horizontal arrow
  $R(f,g) \in V_h$ with $d^-R(f,g)= \cat C$ and $d^+R(f,g)= \cat D$
  with a (op?)-cartesian 2-cell $\alpha \in V_2(R(f,g),R)$ with
  $d^-\alpha = f$ and $d^+\alpha = g$.
\end{definition}

\begin{definition}[Unit]
  A unit $U_{\cat C}\in V_h$ with $d^\pm U_{\cat C} = \cat C$ and an op-cartesian 2-cell
  \[ r : V_h(\cat C, U_{\cat C}) \]
  with $d^\pm r = \id_{\cat C}$.
\end{definition}

\begin{definition}[Tensor]
  If $R,Q \in V_h$ with $d^+R = d^-Q$, a tensor is a horizontal arrow
  $R \odot Q \in V_h$ with $d^+ (R\odot Q) = d^+ R$ and $d^-(R \odot
  Q) = d^- Q$ with an op-cartesian 2-cell
  \[ p : V_h(R,Q;R \odot Q) \]
\end{definition}

\begin{definition}[Hom]

@@

\subsection{Soundness}

\subsection{Completeness}

\begin{definition}
  For any $\Gamma \isadtctx$, we construct a virtual double category
  $V^\Gamma$.
  \begin{enumerate}
  \item $V^\Gamma_v$ is the set of terms $\Gamma \vdash \cat C \isaCat$.
  \item $V^\Gamma_v(\cat C, \cat D)$ is the set of objects $\Gamma
    \pipe \alpha:\cat C \vdash a : \cat D$.
  \item $V_h$ is the set of triples $(\cat C, \cat D, \Gamma \pipe \alpha:\cat C;\beta:\cat D \vdash R \isaSet)$
    where $d^\pm$ are the projections
  \item $V^\Gamma_h(\vec R; S)$ is the set of triples $(a,b,\Gamma
    \pipe \Phi \vdash t : S[a;b])$ where $\Phi$ is the context
    corresponding to $\vec R$ and $d^\pm$ are the obvious projections.

  \item The identity morphisms are $(\cat C, \cat D, R)$ is $\Gamma\pipe\alpha:\cat C, x:R\beta:\cat D \vdash x:R$
  \item Composition $(a,b,t) \circ [(a_1,),\ldots]$ is given by
    substitution in each component: $(a[a_1],b[a_n], t[\phi])$.
  \item associativity and unit rules for composition follow by the
    definition of substitution.
  \end{enumerate}
\end{definition}
